---
title: "Scraping data from social media"
author: "Taken from code prepared by Pablo Barbera"
output: html_document
---


### Scraping web data from Twitter

#### Authenticating

Follow these steps to create your token:

1. Go to apps.twitter.com and sign in.  
2. Click on "Create New App". You will need to have a phone number associated with your account in order to be able to create a token.  
3. Fill name, description, and website (it can be anything, even http://www.google.com). Make sure you leave 'Callback URL' empty.
4. Agree to user conditions.  
5. From the "Keys and Access Tokens" tab, copy consumer key and consumer secret and paste below

```{r, eval=FALSE}
# install.packages("ROAuth")
library(ROAuth)
requestURL <- "https://api.twitter.com/oauth/request_token"
accessURL <- "https://api.twitter.com/oauth/access_token"
authURL <- "https://api.twitter.com/oauth/authorize"
consumerKey <- "XXXXXXXXXXXX"
consumerSecret <- "YYYYYYYYYYYYYYYYYYY"

my_oauth <- OAuthFactory$new(consumerKey=consumerKey,
  consumerSecret=consumerSecret, requestURL=requestURL,
  accessURL=accessURL, authURL=authURL)
```

Run the below line and go to the URL that appears on screen. Then, type the PIN into the console (RStudio sometimes doesn't show what you're typing, but it's there!)

```{r, eval=FALSE}
my_oauth$handshake(cainfo = system.file("CurlSSL", "cacert.pem", package = "RCurl"))
```

Now you can save oauth token for use in future sessions with smappR or streamR. Make sure you save it in a folder where this is the only file.

```{r, eval=FALSE}
save(my_oauth, file="~/git/data-science-workshop/scraping/credentials/twitter-token.Rdata")
```

#### Collecting data from Twitter's Streaming API

Collecting tweets filtering by keyword:

```{r}
#install.packages("streamR")
library(streamR)
load("C:/Users/kevin/Dropbox/credentials/oauth_token27.RData")
filterStream(file.name="blizzard-tweets.json", track="blizzard", 
    timeout=30, oauth=my_oauth)
```

Note the options:
- `file.name` indicates the file in your disk where the tweets will be downloaded  
- `track` is the keyword(s) mentioned in the tweets we want to capture.
- `timeout` is the number of seconds that the connection will remain open  
- `oauth` is the OAuth token we are using

Once it has finished, we can open it in R as a data frame with the `parseTweets` function
```{r}
tweets <- parseTweets("blizzard-tweets.json")
str(tweets)
tweets[1,]
```

And this is how we would capture tweets mentioning multiple keywords:
```{r, eval=FALSE}
filterStream(file.name="snow-tweets.json", 
	track=c("blizzard", "snow", "winter storm", "snowfall"),
    tweets=50, oauth=my_oauth)
```


Finally, it's also possible to collect a random sample of tweets. That's what the "sampleStream" function does:

```{r}
sampleStream(file.name="tweets_random.json", timeout=30, oauth=my_oauth)
```

Here I'm collecting 30 seconds of tweets. And once again, to open the tweets in R...
```{r}
tweets <- parseTweets("tweets_random.json")
```

What is the most retweeted tweet?
```{r}
tweets[which.max(tweets$retweet_count),]
```

What are the most popular hashtags at the moment? We'll use regular expressions to extract hashtags.
```{r}
library(stringr)
ht <- str_extract_all(tweets$text, "#(\\d|\\w)+")
ht <- unlist(ht)
head(sort(table(ht), decreasing = TRUE))
```

How many tweets mention Justin Bieber?
```{r}
length(grep("bieber", tweets$text, ignore.case=TRUE))
```


#### Collecting data from Twitter's REST API


This is how you would extract information from user profiles:

```{r}
library(tweetscores)

reps <- c("RealBenCarson", "tedcruz", "CarlyFiorina", "GrahamBlog", 
    "GovMikeHuckabee", "GovernorPataki", "RandPaul", "marcorubio", 
    "RickSantorum", "bobbyjindal", "GovernorPerry", "realDonaldTrump",
    "JebBush", "GovChristie", "JohnKasich", "ScottWalker", 
    "gov_gilmore")
dems <- c('HillaryClinton', 'SenSanders', "MartinOMalley",
    "LincolnChafee", "JimWebbUSA")
candidates <- c(reps, dems)

users <- getUsersBatch(screen_names=candidates,
                       oauth="C:/Users/kevin/Dropbox/credentials")
str(users)
```

Who is the candidate with the most followers?
```{r}
users[which.max(users$followers_count),]
```

Download up to 3,200 recent tweets from a Twitter account:
```{r}

getTimeline(filename="trump-tweets.json", screen_name="realDonaldTrump", 
    n=1000, oauth="C:/Users/kevin/Dropbox/credentials")
```

What are the most common hashtags?
```{r}
library(streamR)
library(stringr)
tweets <- parseTweets("trump-tweets.json")
ht <- str_extract_all(tweets$text, "#(\\d|\\w)+")
ht <- unlist(ht)
head(sort(table(ht), decreasing = TRUE))
```

Download friends and followers:
```{r}
followers <- getFollowers("kmmunger", 
    oauth="C:/Users/kevin/Dropbox/credentials")
```

What are the most common words that my followers use to describe themselves on Twitter?
```{r, fig.height=6, fig.width=6}
library(quanteda)
# extract profile descriptions
users <- getUsersBatch(ids=followers,
    oauth="C:/Users/kevin/Dropbox/credentials")
# create table with frequency of word use
my_dfm <- dfm(users$description[users$description!=""])

# create wordcloud
par(mar=c(0,0,0,0))
textplot_wordcloud(my_dfm)
```



